---
title: "Dog Breed"
format: html
---

```{r}
breed_traits <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-02-01/breed_traits.csv')
trait_description <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-02-01/trait_description.csv')
breed_rank_all <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-02-01/breed_rank.csv')


```

```{r}
#install.packages("tidymodels")
library(tidymodels)

my_data_splits <- initial_split(breed_traits, prop = 0.5)

exploratory_data <- training(my_data_splits)
test_data <- testing(my_data_splits)
```

Hypothesis: Hunting dogs can be friendly and trained if trained properly like companion dogs.

Below are the top 10 hunting and companion dogs we chose to compare:

-   Hunting dogs

    1.Labrador Retreiver

    2.Shiba Inu

    3.Pointers (German Wirehaired)

    4.American Foxhounds

    5.Spaniels (English Springer)

    6.Black and Tan Coonhounds

    7.Beagles

    8.Brittanys

    9.Norwegian Elkhounds

    10.Spaniel (English Cocker)

-   Companion dogs

    1\. Bichons Frises

    2.Bulldogs

    3\. Dalmatians

    4.Havanese

    5\. Finnish Spitz

    6.Poodles

    7.Pugs

    8.American Eskima Dogs

    9.Boston Terriers

    10.Chinese Crested
